---
title: "Class Project"
format: html
editor: visual
execute: 
  warning: false
---

### Background

This project is attempting to replicate the results found from a multi-year long study in Puerto Rico seeking to examine the levels of various chemicals within various stream testing sites.

This study "Effects of Hurricane Disturbance on Stream Water Concentrations and Fluxes in Eight Tropical Forest Watershed of the Luqillo Experimental Forest, Puerto Rico" (Schaefer et. al)" examined many different sites, but this Quarto doc is limited to recreating a single figure comparing Potassium, Magnesium, Calcium, Ammonia, and Nitric Acid levels at 4 different test sites.

See the subsections below for more detailed explanations of each step of the analysis process.


```{r}
# install and then read in necessary libraries, such as tidyverse, here, zoo, and janitor
source(here::here("scripts", "Cleaning_Code", "install_packages.R"))
source(here::here("scripts", "Cleaning_Code", "read_libraries.R"))
```


### Data
In the data section, we utilized several csv files containing extensive records of sample data from the 4 relevant sites. As part of my developed workflow, I decided to subset out each relevant chemical from each testing site. This was done to help better understand specific trends, keep the data tidy, and ultimately more manageable later in the analysis process. 

```{r}
# stage the sourced and cleaned data from the the "data_cleaning_source.R" script. This contains the subsets of each testing site by chemical. This is done so that each set may 
source(here::here("scripts", "Cleaning_Code", "data_cleaning_source.R"))
```

### Methods
The main function utilized for this analysis was a function titled "rolling_average_script" which was replication of a shared within class. The principle difference between the two is that, to account for Na results within the data sets, a small edit was made when calculating the mean. See the code chunk below for the details.

```{r}

rolling_average_script <- function(focal_date, dates, conc, win_size_wks) {
  is_in_window <- (dates > focal_date - (win_size_wks / 2) * 7) &
    (dates < focal_date + (win_size_wks / 2) * 7)
  window_conc <- conc[is_in_window]
  result <- mean(window_conc, na.rm = TRUE)
  return(result)
}

```

We then utilized the above function within a different R script, and applied it to the many different subsets. To keep this document more readable, we've sourced in that script. 

```{r}
source(here::here("scripts", "Analysis_Scripts", "staged_analysis.R"))
```

After the rolling average function was applied to each of the location/chemical subsets, we then rejoined the data frames by their chemical sample for visualization.

```{r}
joined_sites_K <- bind_rows(BQ1_K, BQ2_K, BQ3_K, PRM_K)
joined_sites_NO3 <- bind_rows(BQ1_NO3, BQ2_NO3, BQ3_NO3, PRM_NO3)
joined_sites_Mg <- bind_rows(BQ1_Mg, BQ2_Mg, BQ3_Mg, PRM_Mg)
joined_sites_NH4 <- bind_rows(BQ1_NH4, BQ2_NH4, BQ3_NH4, PRM_NH4)
joined_sites_Ca <- bind_rows(BQ1_Ca, BQ2_Ca, BQ3_Ca, PRM_Ca)
```


### Results
Finally, we visualized the sample data trends over the period from the sites' earliest recordings through the end of 1995 to match the graphs in the original figure. Due to the methodology utilized, we were unable to facet a single data frame to include the 5 graphs below as a single joined plot. Were  I to repeat this analysis, I would try to keep all the data sets in one table, and then conduct the analysis there, as it would likely allow for a better visualization of the data.  

EDIT* after instructor feedback, I utilized functionality from the patchwork library to stack the 5 graphs in one column. keeping the original 5 graphs for comparison and consistency with previous text. 

```{r}
#| echo: false
Potassium_plot <- ggplot() +
  geom_line(data = joined_sites_K, aes(x = Sample_Date, y = rolling_average, color = Sample_ID)) + 
  labs(x = "Years",
       y = "K mg l-1",
       color = " ")

Nitrate_plot <- ggplot() +
  geom_line(data = joined_sites_NO3, aes(x = Sample_Date, y = rolling_average, color = Sample_ID)) + 
  labs(x = "Years",
       y = "NO3-N ug l-1",
       color = " ")

MG_plot <- ggplot() +
  geom_line(data = joined_sites_Mg, aes(x = Sample_Date, y = rolling_average, color = Sample_ID)) + 
  labs(x = "Years",
       y = "Mg mg l-1")

Ca_plot <- ggplot() +
  geom_line(data = joined_sites_Ca, aes(x = Sample_Date, y = rolling_average, color = Sample_ID)) + 
  labs(x = "Years",
       y = "Ca mg l-1",
       color = " ")

Ammonia_plot <- ggplot() +
  geom_line(data = joined_sites_NH4, aes(x = Sample_Date, y = rolling_average, colour = Sample_ID)) + 
  labs(x = "Years",
       y = "NH4-N ug l-1",
       color = " ")

Nitrate_plot / Potassium_plot / MG_plot / Ca_plot / Ammonia_plot
```



## Potassium Plot

```{r}
ggplot() +
  geom_line(data = joined_sites_K, aes(x = Sample_Date, y = rolling_average, colour = Sample_ID)) + 
  labs(x = "Years",
       y = "K mg l-1",
       color = " ")
  

```

## Nitrate Plot

```{r}
ggplot() +
  geom_line(data = joined_sites_NO3, aes(x = Sample_Date, y = rolling_average, colour = Sample_ID)) + 
  labs(x = "Years",
       y = "NO3-N ug l-1",
       color = " ")

```

## Magnesium Plot

```{r}
ggplot() +
  geom_line(data = joined_sites_Mg, aes(x = Sample_Date, y = rolling_average, colour = Sample_ID)) + 
  labs(x = "Years",
       y = "Mg mg l-1",
       color = " ")
```

## Calcium Plot

```{r}
ggplot() +
  geom_line(data = joined_sites_Ca, aes(x = Sample_Date, y = rolling_average, colour = Sample_ID)) + 
  labs(x = "Years",
       y = "Ca mg l-1",
       color = " ")
```

## Ammonia Plot

```{r}
ggplot() +
  geom_line(data = joined_sites_NH4, aes(x = Sample_Date, y = rolling_average, colour = Sample_ID)) + 
  labs(x = "Years",
       y = "NH4-N ug l-1",
       color = " ")
```


## Acknowledgements

Schaefer DouglasA, McDowell WH, Scatena FN, Asbury CE. "Effects of hurricane disturbance on stream water concentrations and fluxes in eight tropical forest watersheds of the Luquillo Experimental Forest, Puerto Rico." Journal of Tropical Ecology. 2000;16(2):189-207. <doi:10.1017/S0266467400001358>